Question 1
==============================================================

__Goal__: estimate the average number of steps taken per day

Written by [L. Collado-Torres](http://bit.ly/13MBoy8).

To make it easier for me to write this document, the interpretation of each plot is included as comments in the R code. So toggle the R code to view the comments and interpretation if you are curious.

# Setup

```{r setup}
## Packages used
library(fitbitR)
library(xts)
library(forecast)

## Load the data
data <- preprocess(fitbitData)

## Complete predictions
types <- c("lm", "poisson", "means", "overall-mean")
datap <- lapply(types, function(x) { fitbitPred(data, method=x)})

## All data
all <- c(list(data), datap)
names(all) <- c("original", types)

```

# Simple mean

Below I take the sample mean and deviation after binning by day.

```{r simplest}
## This is the simplest solution although it might have some problems.
simpleM <- function(data) {
	mean(tapply(data$nSteps, data$Date, sum), na.rm=TRUE)
}
simpleSE <- function(data) {
	sd(tapply(data$nSteps, data$Date, sum), na.rm=TRUE) / sqrt(length(unique(data$Date)))
}
simple.m <- unlist(lapply(all, simpleM))
simple.se <- unlist(lapply(all, simpleSE))
simple <- data.frame(simple.m, simple.se)
simple

```

Here it is important to note that the sd decreases when using predicted data. The estimate however is higher for prediction methods lm, poisson and means. 

# Binning by day

```{r binday}

## Summarize the information by day. Drops interval and time. The binDay function is included in fitbitR:
# binDay
binned <- lapply(all, binDay)

## Quickly check how the distribution looks now that the data has been binned by day.
xx <- lapply(1:5, function(i) { 
	x <- binned[[i]]
	hist(x$nSteps, col="light blue", main=paste("Prediction type:", names(binned)[i]), xlab="Number of steps by day", ylim=c(0, 35))
	return(invisible())
})

## Are the means different due to the predictions?
for(i in 2:5) {
	print(t.test(binned[[1]]$nSteps, binned[[i]]$nSteps))
}
## No significative difference seen with t-tests.

```


# ARIMA binned by day

```{r arimaBin}
arimaE <- function(data, type, order="Time") {
	datac <- data[complete.cases(data),]
	ts.c <- xts(datac$nSteps, order.by=datac[, order])
	
	## Autocorrelation plot
	acf(ts.c, main=type)
	
	## Fit arima model
	fit.a <- arima(ts.c)
	
	return(fit.a)
	
}
```

## ARIMA(0, 0, 0)

Below are the ACF plots and model summaries when using ARIMA(0, 0, 0) on the data binned by day.

```{r arimaBin2}
## ARIMA with the binned data
binari <- lapply(names(binned), function(x) { arimaE(binned[[x]], x, order="Date") })
names(binari) <- names(binned)
binari
```

The coefficients are the same using the binned data vs the simple method. However, the SD is smaller.

```{r arimaBin3}
## The coefficients are the same using the binned data vs the simple method. However, the SD is smaller.
binari.sum <- data.frame("Coef" = unlist(lapply(binari, coef)), "SE" = unlist(lapply(binari, function(x) sqrt(x$var.coef["intercept", "intercept"]))))
binari.sum
```

## Auto ARIMA: (0, 0, 1)

Lets try with the forecast package in order to use auto.arima for selecting the ARIMA models.

```{r forecast}

## Similar to arimaE function but using auto.arima
arimaF <- function(data, order="Time") {
	datac <- data[complete.cases(data),]
	ts.c <- xts(datac$nSteps, order.by=datac[, order])
	
	## Fit arima model
	fit.a <- auto.arima(ts.c)
	
	return(fit.a)
}

```

Auto ARIMA results.

```{r forecast2}
## Fit the auto-arima. Note that all models ended up as ARIMA(0, 0, 1) which is equivalent to MA(1)
binariauto <- lapply(binned, arimaF, order="Date")
binariauto
```

Summary information.

```{r forecast3}
## Summarize the information
binariauto.sum <- data.frame("Coef" = unlist(lapply(binariauto, function(x){ coef(x)["intercept"]})), "SE" = unlist(lapply(binariauto, function(x) sqrt(x$var.coef["intercept", "intercept"]))))
binariauto.sum
```


# ARIMA not binned

Fit ARIMA model with automatic selection of parameters without binning the data by day.

## ARIMA (0, 0, 0)

ARIMA (0, 0, 0) results.

```{r notBin1}
arimed <- lapply(names(all), function(x) { arimaE(all[[x]], x) })
names(arimed) <- names(all)
arimed
```

Summary information.

```{r notBin1sum}
## Summarize the information
arimed.sum <- data.frame("Coef" = unlist(lapply(arimed, function(x){ coef(x)["intercept"]})) * 288, "SE" = unlist(lapply(arimed, function(x) sqrt(x$var.coef["intercept", "intercept"]))) * 288 )
arimed.sum
```


## Auto ARIMA: (3, 0, 3)

Auto ARIMA results.

```{r notBin2}
## Fitting with auto-arima. Models were either ARIMA(2, 0, 3) or ARIMA(3, 0, 3).
arimed2 <- lapply(all, arimaF)
arimed2
```

Summary information.

```{r notBin2sum}
## Summarize the information
arimed2.sum <- data.frame("Coef" = unlist(lapply(arimed2, function(x){ coef(x)["intercept"]})) * 288, "SE" = unlist(lapply(arimed2, function(x) sqrt(x$var.coef["intercept", "intercept"]))) * 288 )
arimed2.sum
```




# Summary

```{r summary}
fullSummary <- cbind(simple, binari.sum, binariauto.sum, arimed.sum, arimed2.sum)
colnames(fullSummary) <- paste0(rep(c("Simple", "Binned", "BinnedAuto", "NotBin", "NotBinAuto"), each=2), rep(c("Coef", "SE"), 5))
fullSummary
```


* Using the predicted information does not change the estimate significantly as verified using t-tests with the data binned by day. 
* Estimating via sample mean and the SE leads to lower SE values. This could be because of not taking into account that the data are correlated: what you did day 1 can affect day 2.
* Using an ARIMA(0, 0, 0) on the data binned by day leads to the same estimates as the simple method but with larger SEs.
* Using an ARIMA(0, 0, 1) (equivalent to MA(1)) on binned by day data we can get estimates (very similar to the _simple_ ones) which take into account some of the information from the previous day (MA estimate). The SE goes up though.
* Using an ARIMA(0, 0, 0) on the data not binned by day leads to the same estimates as the simple method and smaller SEs.
* Using an ARIMA(3, 0, 3) on the data not binned by day we get slightly different estimates with the largest SE of all methods. The smaller SE in the other methods could be due to under-estimating the variability in the mean number of steps per interval (per day if using binned data).

# Using q1()

Now I can get the estimate using either the means or the auto.arima model (without binning by day) using q1().

```{r q1fun}
## Using the mean method
q1.mean <- lapply(all, q1, method="mean")
do.call(rbind, q1.mean)

## Using the auto.arima method
q1.aa <- lapply(all, function(x) { y <- q1(x, method="auto.arima", acf=FALSE); y$Estimate })
do.call(rbind, q1.aa)
```


# Reproducibility

This report was generated using [knitrBootstrap](https://github.com/jimhester/knitrBootstrap).

```{r reproducibility}
sessionInfo()
print(proc.time())
```


